---
title: "simulation results for kamplite"
author: "Julia Wrobel"
date: '`r format(Sys.time(), "%Y-%m-%d")`'
output:
  html_document: 
    code_folding: hide
    toc: true
    toc_float: true
hitheme: tomorrow
highlighter: highlight.js
---

# Overview

This file loads data objects stored from simulations and produces figures for the supplement, specifically related to varying thinning probabilities across different simulation settings for kamplite.  Libraries and code are sourced below. 

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(viridis)
library(patchwork)
library(latex2exp)

knitr::opts_chunk$set(echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.width = 9,
  fig.height = 4,
  fig.path = '../output/'
)

theme_set(theme_bw() + theme(legend.position = "bottom"))

# colorblind-safe palette for thinning probability (Okabe-Ito subset), chosen
# to avoid the hues used for K/Kinhom/perm/KAMP/KAMP lite in the other figures
# (purple #6A3D9A, blue #1F78B4, green #4DAF4A, vermillion #D55E00, orange #E69F00)
thinning_colors <- c(
  "0"    = "#000000",  # black
  "0.25" = "#56B4E9",  # sky blue
  "0.5"  = "#F0E442",  # yellow
  "0.75" = "#CC79A7"   # reddish purple
)
```

# Source code

Source file for consolidating simulations

```{r}
source(here::here("source", "utils.R"))
```

# Univariate

## Clean and process simulation data

Here we load and process simulation results. Full simulations involve 36 scenarios with 50 simulated datasets each.


```{r}
# load only results for VB and TS
files = list.files(here::here("output", "kamplite", "20260901"), 
									 full.names =  TRUE)

# call merge_files() defined in utils.R
df = map_dfr(files, merge_files)

```




```{r}
# make comp times dataset
times_df = df %>%
  filter(lambda_n != 2000) %>%
  select(r, iter, type, lambda_n,
         abundance,
         p_thin,
         time) %>%
  filter(r == .5) %>%
  arrange(p_thin)

df = df %>%
  filter((type %in% c("hom", "inhom") & r == 0.5) | (type %in% c("homClust", "inhomClust") & r == 1.25),
         lambda_n != 2000
         ) %>%
  mutate(abundance = factor(abundance),
         value = trans-csr)  %>%
  #filter(value < 25) %>%
    mutate(type = factor(type, levels = c("hom", "inhom", "homClust", "inhomClust"),
                            labels = c("homogeneous",
                                       "inhomogeneous",
                                       "homogeneous + clustered",
                                       "inhomogeneous + clustered")),
         lambda_n = factor(lambda_n)) 
```



## Figures 


```{r}
 p_time = times_df %>%
  filter(
      type == "inhomClust",
         p_thin != 0.9,
         #abundance != 0.001
         ) %>%
  group_by(p_thin, lambda_n, abundance) %>%
  summarize(time = median(time)) %>%
  ungroup() %>%
  mutate(Thinning = factor(p_thin),
         lambda_n = factor(lambda_n)) %>%
  ggplot(aes(lambda_n, time, color = Thinning, 
             group = interaction(lambda_n, Thinning))) +
  geom_point(size = 2) +
  geom_line(aes(group = Thinning), linewidth = 1.25) +
  theme_minimal() +
  theme(legend.position =c(0.85, 0.6)) +
  geom_line(aes(group = Thinning), linewidth = 1.25) +
  geom_hline(yintercept = log(1), linetype = 4, color = "gray") +
  scale_color_manual(values = thinning_colors) +
  labs(x = expression(lambda[n]),
       y = "Median time in seconds",
       title = "Univariate KAMP lite: computation time") +
   facet_wrap(~abundance, ncol = 4, scales= "free_x", labeller = label_both)


p_time
```



Power = P(reject Ho|H1 True)

```{r}
power_df = df %>%
    filter(type %in% c("inhomogeneous + clustered"),
           r == 1.25,
           p_thin != 0.9,
           ) %>%
  group_by(type, p_thin, lambda_n, abundance) %>%
  summarize(power = sum(pvalue<= 0.05)/n(),
            n_sim = n()) %>%
  ungroup() %>%
    mutate(abundance = factor(abundance),
         lambda_n = factor(lambda_n),
         p_thin = factor(p_thin))

p_power = power_df %>%
  ggplot(aes(lambda_n, power, group = p_thin,
             color = p_thin, linetype = p_thin)) +
  geom_point(size = 2) +
  geom_line(linewidth = 1.25) +
  labs(y = "Power", x = expression(lambda[n]),
       title = "Univariate KAMP lite: power") +
  #ylim(0.025, 1) +
  geom_hline(yintercept = 0.9, linetype = 4) +
  theme_minimal() +
  theme(legend.position = "none", legend.title = element_blank()) +
  scale_color_manual(values = thinning_colors) +
  facet_wrap(~abundance, ncol = 4, scales= "free_x", labeller = label_both)




```


# bivariate


```{r}
# load only results for VB and TS
files = list.files(here::here("output", "kamplite_bivariate", "20260909"),
									 full.names =  TRUE)

# call merge_files() defined in utils.R
df = map_dfr(files, merge_files)

```




```{r}
df = df %>%
  filter((type == "inhom" & r == 0.5) | (type == "inhomClust" & r == 1.25)
         ) %>%
  mutate(abundance1 = factor(abundance1),
         abundance2 = factor(abundance2),
         value = trans-csr)  %>%
  #filter(value < 25) %>%
    mutate(type = factor(type, levels = c("inhom", "inhomClust"),
                            labels = c("inhomogeneous",
                                       "inhomogeneous + clustered")),
         lambda_n = factor(lambda_n))
```



## Figures



Power = P(reject Ho|H1 True)

```{r}
power_df = df %>%
    filter(type %in% c("inhomogeneous + clustered"),
           r == 1.25,
           p_thin != 0.9
           ) %>%
  rename(abundance = abundance1) %>%
  group_by(type, p_thin, lambda_n, abundance, abundance2) %>%
  summarize(power = sum(pvalue<= 0.05)/n(),
            n_sim = n()) %>%
  ungroup() %>%
    mutate(lambda_n = factor(lambda_n),
         p_thin = factor(p_thin))

p_power_biv = power_df %>%
  ggplot(aes(lambda_n, power, group = p_thin,
             color = p_thin)) +
  geom_point(size = 2) +
  geom_line(aes(linetype = p_thin), linewidth = 1.25) +
  labs(y = "Power", x = expression(lambda[n]),
       title = "Bivariate KAMP lite: power") +
  geom_hline(yintercept = 0.9, linetype = 4) +
  theme_minimal() +
  theme(legend.position = "none") +
  scale_color_manual(values = thinning_colors) +
  facet_wrap(~abundance, labeller = label_both)


```






```{r supp_fig_kamplite_power, fig.height = 9}
p_time /p_power/p_power_biv
```



