---
title: "simulation results for kamplite (bivariate)"
author: "Julia Wrobel"
date: '`r format(Sys.time(), "%Y-%m-%d")`'
output:
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
hitheme: tomorrow
highlighter: highlight.js
---

# Overview

This file loads data objects stored from simulations and produces figures for the supplement, specifically related to varying thinning probabilities across different simulation settings for kamplite, bivariate (immune1 x immune2) setting. Libraries and code are sourced below.

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(viridis)
library(patchwork)
library(latex2exp)

knitr::opts_chunk$set(echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.width = 9,
  fig.height = 4,
  fig.path = '../output/'
)

theme_set(theme_bw() + theme(legend.position = "bottom"))
```

# Source code

Source file for consolidating simulations

```{r}
source(here::here("source", "utils.R"))
```


# Expectation

## Clean and process simulation data

Here we load and process simulation results. Full simulations involve 54 scenarios with 50 simulated datasets each.


```{r}
# load only results for VB and TS
files = list.files(here::here("output", "kamplite_bivariate", "20260908"),
									 full.names =  TRUE)

# call merge_files() defined in utils.R
df = map_dfr(files, merge_files)

```




```{r}
# make comp times dataset
times_df = df %>%
  select(r, iter, type, lambda_n,
         abundance1, abundance2,
         p_thin,
         time) %>%
  filter(r == .5) %>%
  arrange(p_thin) %>%
  mutate(abundance1 = factor(abundance1),
         abundance2 = factor(abundance2))

df = df %>%
  filter((type == "inhom" & r == 0.5) | (type == "inhomClust" & r == 1.25)
         ) %>%
  mutate(abundance1 = factor(abundance1),
         abundance2 = factor(abundance2),
         value = trans-csr)  %>%
  #filter(value < 25) %>%
    mutate(type = factor(type, levels = c("inhom", "inhomClust"),
                            labels = c("inhomogeneous",
                                       "inhomogeneous + clustered")),
         lambda_n = factor(lambda_n))
```



## Figures


```{r}
 p_time = times_df %>%
  filter(type == "inhom",
         p_thin != 0.9,
         abundance2 == 0.1
         ) %>%
  group_by(p_thin, lambda_n, abundance1, abundance2) %>%
  summarize(time = median(time)) %>%
  ungroup() %>%
  mutate(Thinning = factor(p_thin),
         lambda_n = factor(lambda_n, levels = c(5000, 10000, 20000))) %>%
  ggplot(aes(lambda_n, time, color = Thinning,
             group = Thinning)) +
  geom_point() +
  geom_line(linewidth = 1.25) +
  theme_minimal() +
  theme(legend.position = c(0.75, 0.6)) +
  geom_hline(yintercept = log(1), linetype = 4, color = "gray") + 
  facet_wrap(~abundance1, labeller = label_both)
  

p_time

```



Power = P(reject Ho|H1 True)

```{r}
power_df = df %>%
    filter(type %in% c("inhomogeneous + clustered"),
           r == 1.25,
           p_thin != 0.9,
          abundance2 == 0.01
           ) %>%
  group_by(type, p_thin, lambda_n, abundance1, abundance2) %>%
  summarize(power = sum(pvalue<= 0.05)/n(),
            n_sim = n()) %>%
  ungroup() %>%
    mutate(lambda_n = factor(lambda_n),
         p_thin = factor(p_thin))

p_power = power_df %>%
  ggplot(aes(lambda_n, power, group = p_thin,
             color = p_thin)) +
  geom_point(size = 3) +
  geom_line(linewidth = 1.25) +
  labs(y = "Power", x = expression(lambda[n]),
       title = "Power") +
  #ylim(0.025, 1) +
  geom_hline(yintercept = 0.9, linetype = 4) +
  theme_minimal() +
  theme(legend.position = "bottom") +
  facet_wrap(~abundance1, labeller = label_both)

p_power


```




```{r fig_kamplite_bivariate_times, fig.height = 6}
p_time /p_power
```
